---
title: "Use Case D: Investigating changes in population mental health related to COVID-19"
author: ""
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE,
                      out.width = "100%")
library(tidyverse)
library(readr)
library(ggplot2)
#library(MASS)
library(dplyr) # if the subgroups are unique
library(lubridate)
library(plotly)
library(DT)

warning = function(x, options) {
     paste('\n\n<div class="alert alert-warning">',
           gsub('##', '\n', gsub('^##\ Warning:', '**Warning**', x)),
           '</div>', sep = '\n')
}


calculate_socecon_lvl_ratio <- function(p1) {
  
  if(!("5" %in% colnames(p1)))
  {
    p1$`5` <- NA
  }
  if(!("4" %in% colnames(p1)))
  {
    p1$`4` <- NA
  }
  if(!("3" %in% colnames(p1)))
  {
    p1$`3` <- NA
  }
  if(!("2" %in% colnames(p1)))
  {
    p1$`2` <- NA
  }
  if(!("1" %in% colnames(p1)))
  {
    p1$`1` <- NA
  }
  p1 <- p1 %>% mutate(ratio = `1` /  sum(c(`5`,`4`,`3`,`2`),  na.rm=T))
  return(p1)
}


```

## Use Case D on utilization of health care services by mental health patients during the COVID pandemic ##
The COVID-19 pandemic and the public health measures imposed to tackle the surge of cases has imposed a heavy burden in population mental health as people has limited social contact or has suffer from the consequences of disease and the derived economic downturn. This use case will explore changes in the utilization of health care services by people with depression or ansiety (as most prevalent mental health issues) derived as indirect effects of the COVID-19 pandemic at population level.   


### LOCAL ANALYSES ###
This analyses corresponds with the local part of the analysis (country-specific). 
Please, provide feedback on your outputs, both data quality analysis (EDA), HTML report and aggregated output to the Use Case B leaders to complete the overall analyses.

</br>

```{r summary_proc , message=FALSE, echo=FALSE, warning=FALSE, error=FALSE}
## Read your data
ucd <- read_delim("usecaseD_synthetic.csv", delim = "|", escape_double = FALSE, trim_ws = TRUE)

ucd <- ucd %>% mutate(age_group = factor(cut(age_nm, seq(0, 109, 5), right = FALSE, 
                        labels = c("0-4", "5-9", "10-14", "15-19", "20-24", 
                                   "25-29", "30-34", "35-39", "40-44", "45-49",
                                   "50-54", "55-59", "60-64", "65-69", "70-74",
                                   "75-79", "80-84", "85-89", "90-94", "95-99", 
                                   "99+")))) 

#Create depression and anxiety diagnosis from ICD codes
ucd$dep = 0
ucd$anx = 0

#ICD-10_mc
ucd$dep<-as.numeric(substr(ucd$diagnosis, 1,2)=="F3")
ucd$anx<-as.numeric(substr(ucd$diagnosis, 1,2)=="F4")

#ICD-9_mc
ucd$dep[ucd$dep == 0] <-as.numeric(ucd$diagnosis[ucd$dep == 0] %in% c("296.20","296.22","296.23","296.30","296.32","296.33","300.4","311")) 

ucd$anx[ucd$anx == 0] <-as.numeric(ucd$diagnosis[ucd$anx == 0] %in% c("300.00","300.01","300.02","300.09")) 



#SNOMED
ucd$dep[ucd$dep == 0] <-as.numeric(ucd$diagnosis[ucd$dep == 0] %in% c("46206005","370143000","712823008")) 
ucd$anx[ucd$anx == 0] <-as.numeric(ucd$diagnosis[ucd$anx == 0] %in% c("197480006","21897009")) 


init_year <- year(min(min(ucd$dx_date,na.rm = TRUE),min(ucd$drug_date,na.rm = TRUE),min(ucd$visit_date,na.rm = TRUE)))
end_year <- year(max(max(ucd$dx_date,na.rm = TRUE),max(ucd$drug_date,na.rm = TRUE),max(ucd$visit_date,na.rm = TRUE)))
diff_years <- (end_year - init_year) + 1


year_months_0 <- data.frame(month=rep(1:12,diff_years),year= rep(init_year:end_year,each=12),sex = 0)
year_months_1 <- data.frame(month=rep(1:12,diff_years),year= rep(init_year:end_year,each=12),sex = 1)
year_months <- rbind(year_months_0, year_months_1)
rownames(year_months) <- NULL



dx_per_month_year <- ucd %>% 
    mutate(month = month(dx_date), year = year(dx_date)) %>% dplyr::select(year,month,sex) %>% 
    group_by(year, month, sex) %>% count() 
names(dx_per_month_year) <- c("year","month","sex","total_dx")



dx_per_month_year <- merge(year_months, dx_per_month_year, by=c("year","month","sex"),all.x = TRUE)
dx_per_month_year[is.na(dx_per_month_year)] <- 0


drug_per_month_year <- ucd %>% 
  mutate(month = month(drug_date), year = year(drug_date)) %>% dplyr::select(year,month,sex) %>% 
  group_by(year, month, sex) %>% count() 
names(drug_per_month_year) <- c("year","month","sex","total_drug")

drug_per_month_year <- merge(year_months, drug_per_month_year, by=c("year","month","sex"),all.x = TRUE)
drug_per_month_year[is.na(drug_per_month_year)] <- 0



dep_mo <- ucd %>% filter(dep == 1) %>%  mutate(month = month(dx_date), year = year(dx_date)) %>%
  dplyr::select(year,month,sex,dep) %>% group_by(year, month, sex) %>% count() 
names(dep_mo) <- c("year","month","sex","total_dep")  
  

dep_mo <- merge(year_months, dep_mo, by=c("year","month","sex"),all.x = TRUE)
dep_mo[is.na(dep_mo)] <- 0


anx_mo <- ucd %>% filter(anx == 1) %>%  mutate(month = month(dx_date), year = year(dx_date)) %>% dplyr::select(year,month,sex,anx) %>%  
  group_by(year, month, sex) %>% count() 

names(anx_mo) <- c("year","month","sex","total_anx")  


anx_mo <- merge(year_months, anx_mo, by=c("year","month","sex"),all.x = TRUE)
anx_mo[is.na(anx_mo)] <- 0

prim_visit_mo <- ucd %>% mutate(month = month(visit_date), year = year(visit_date)) %>% dplyr::select(year,month,sex,prim_visit) %>% na.omit(.) %>%
  group_by(year, month, sex) %>% 
  summarise(total_prim_visit = sum(prim_visit))



prim_visit_mo <- merge(year_months, prim_visit_mo, by=c("year","month","sex"),all.x = TRUE)
prim_visit_mo[is.na(prim_visit_mo)] <- 0

hosp_visit_mo <- ucd %>% mutate(month = month(visit_date), year = year(visit_date)) %>% dplyr::select(year,month,sex,hosp_visit) %>% na.omit(.) %>% 
  group_by(year, month, sex) %>% 
  summarise(total_hosp_visit = sum(hosp_visit))
  


hosp_visit_mo <- merge(year_months, hosp_visit_mo, by=c("year","month","sex"),all.x = TRUE)
hosp_visit_mo[is.na(hosp_visit_mo)] <- 0
  
emer_visit_mo <- ucd %>% mutate(month = month(visit_date), year = year(visit_date)) %>% dplyr::select(year,month,sex,emer_visit) %>% na.omit(.) %>%
  group_by(year, month, sex) %>% 
  summarise(total_emer_visit = sum(emer_visit))


emer_visit_mo <- merge(year_months, emer_visit_mo, by=c("year","month","sex"),all.x = TRUE)
emer_visit_mo[is.na(emer_visit_mo)] <- 0


summary_table <- merge(dx_per_month_year, drug_per_month_year, by=c("year","month","sex"),all.x = TRUE)
summary_table <- merge(summary_table, dep_mo, by=c("year","month","sex"),all.x = TRUE)
summary_table <- merge(summary_table, anx_mo, by=c("year","month","sex"),all.x = TRUE)
summary_table <- merge(summary_table, prim_visit_mo, by=c("year","month","sex"),all.x = TRUE)
summary_table <- merge(summary_table, hosp_visit_mo, by=c("year","month","sex"),all.x = TRUE)
summary_table <- merge(summary_table, emer_visit_mo, by=c("year","month","sex"),all.x = TRUE)
summary_table <- summary_table %>% filter(year >= 2017)

if(class(summary_table$total_dx)=="list"){summary_table$total_dx <- 0}
if(class(summary_table$total_drug)=="list"){summary_table$total_drug <- 0}
if(class(summary_table$total_dep)=="list"){summary_table$total_dep <- 0}
if(class(summary_table$total_anx)=="list"){summary_table$total_anx <- 0}
if(class(summary_table$total_prim_visit)=="list"){summary_table$total_prim_visit <- 0}
if(class(summary_table$total_hosp_visit)=="list"){summary_table$total_hosp_visit <- 0}
if(class(summary_table$total_emer_visit)=="list"){summary_table$total_emer_visit <- 0}



dataset1 <- ucd %>% filter(!is.na(dx_date)) %>%
  mutate(month = month(dx_date), year = year(dx_date)) %>%  
  count(year,month,age_group,sex,socecon_lvl_cd) %>% rename(total_dx = n )


dataset2 <- ucd %>% filter(!is.na(visit_date)) %>%
  mutate(month = month(visit_date), year = year(visit_date)) %>%  
  count(year,month,age_group,sex,socecon_lvl_cd) %>% rename(total_drug = n )


dataset3 <- ucd%>% filter(!is.na(visit_date)) %>%
  mutate(month = month(visit_date), year = year(visit_date)) %>% 
  group_by(year,month,age_group,sex,socecon_lvl_cd) %>%
  summarise(total_prim_visit = sum(prim_visit))

dataset4 <- ucd%>% filter(!is.na(visit_date)) %>%
  mutate(month = month(visit_date), year = year(visit_date)) %>% 
  group_by(year,month,age_group,sex,socecon_lvl_cd) %>%
  summarise(total_hosp_visit = sum(hosp_visit))

dataset5 <- ucd%>% filter(!is.na(visit_date)) %>%
  mutate(month = month(visit_date), year = year(visit_date)) %>% 
  group_by(year,month,age_group,sex,socecon_lvl_cd) %>%
  summarise(total_emer_visit = sum(emer_visit))

summary_table_aggregated <- merge(x=dataset1,y=dataset2,by=c("year","month","age_group","sex","socecon_lvl_cd"),all=TRUE)
summary_table_aggregated <- merge(x=summary_table_aggregated,y=dataset3,by=c("year","month","age_group","sex","socecon_lvl_cd"),all=TRUE)
summary_table_aggregated <- merge(x=summary_table_aggregated,y=dataset4,by=c("year","month","age_group","sex","socecon_lvl_cd"),all=TRUE)
summary_table_aggregated <- merge(x=summary_table_aggregated,y=dataset5,by=c("year","month","age_group","sex","socecon_lvl_cd"),all=TRUE)
summary_table_aggregated[is.na(summary_table_aggregated)] <- 0
summary_table_aggregated <- summary_table_aggregated %>% filter(year >= 2017)



```



```{r plot1 , message=FALSE, echo=FALSE, warning=FALSE, error=FALSE}
# convert values to months
month_breaks <- c(1,2,3,4,5,6,7,8,9,10,11,12)
month_names <-
  c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec")




#############################################################################
#                             TOTAL _DX                                     #
#############################################################################

summary_table$sex[summary_table$sex == 0] <- "men"
summary_table$sex[summary_table$sex == 1] <- "women"


p1<- ggplot(summary_table, aes(x=month, y=total_dx, color = as.factor(year), linetype = as.factor(sex), fill = as.factor(sex))) +
  geom_line(size = 1) +
  theme(panel.grid.major.x = element_line( color = "grey"),
        panel.grid.major.y = element_blank(), panel.border = element_blank(),
        panel.grid.minor = element_blank(), panel.background = element_blank(),
        axis.line = element_line(color = "black")) +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_vline(xintercept = 3.76, color = "red", linetype = 2, size = 1) +
  labs(subtitle = paste0("n = ", sum(summary_table$total_dx)),
       title = "All events",
       color = "Year",
       linetype = "Sex") +
  ylab(paste0("Absolute number of cases (N=", sum(summary_table$total_dx),")"))  +
  expand_limits(y = 0) +
  scale_x_continuous(name = "Month", breaks = month_breaks[1:12], labels = month_names, expand = c(0,0)) +
  theme(plot.subtitle = element_text(hjust = 1), plot.caption = element_text(hjust = 1))

ggplotly(p1)


```

```{r ratio, message=FALSE, echo=FALSE, warning=FALSE, error=FALSE}

highest_socecon_lvl <- length(levels(factor(ucd$socecon_lvl_cd)))
highest_socecon_lvl <- `highest_socecon_lvl`

dataAva <- nrow(ucd %>% 
  mutate(month = str_pad(month(dx_date), 2, pad = "0"), year = year(dx_date)) %>% filter(year >= 2017) %>% mutate(year_key = paste0(year,'-',month)) %>%
  group_by(year_key) %>%
  count(socecon_lvl_cd)) >0

if(!is.na(highest_socecon_lvl) && highest_socecon_lvl!=0 && dataAva){
p1 <-
  ucd %>% 
  mutate(month = str_pad(month(dx_date), 2, pad = "0"), year = year(dx_date)) %>% filter(year >= 2017) %>% mutate(year_key = paste0(year,'-',month)) %>%
  group_by(year_key) %>%
  count(socecon_lvl_cd) %>%
  pivot_wider(id_cols = year_key, names_from = socecon_lvl_cd, values_from = n)

p1 <- calculate_socecon_lvl_ratio(p1) 
  
p1 <- p1 %>% 
  ggplot(aes(x= year_key, y=ratio, group = 1)) +
  geom_line(size=1) +
  geom_smooth(method = "loess",show.legend=TRUE) + 
  theme(panel.grid.major.x = element_line( color = "grey"),
        panel.grid.major.y = element_blank(), panel.border = element_blank(),
        panel.grid.minor = element_blank(), panel.background = element_blank(),
        axis.line = element_line(color = "black")) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Ratio of most to least deprived",
       y= "Ratio most deprived / any socioeconomic level",
       x= "Time",
       color = "Year") +
  theme(plot.subtitle = element_text(hjust = 1), plot.caption = element_text(hjust = 1)) 



ggplotly(p1) %>%
  layout(title = list(text = paste0('Ratio of most to least deprived',
                                    '<br>',
                                    '<sup>',
                                    '*Loess regression',
                                    '</sup>')))
}


```

`r if(!dataAva){ warning("Warning: No information available on 'dx_date'") }`


`r if(is.na(highest_socecon_lvl) || highest_socecon_lvl==0 ){ warning("Warning: No information available on 'socecon_lvl_cd'") }`



</br>
</br>

```{r plot2,  message=FALSE, echo=FALSE, warning=FALSE, error=FALSE}
summary_table$sex[summary_table$sex == "men"] <- 0
summary_table$sex[summary_table$sex == "women"] <- 1

p2<-summary_table %>%  filter(sex==1) %>% ggplot(aes(x=month, y=total_dx, color = as.factor(year))) +
  geom_line(size = 1) +
  theme(panel.grid.major.x = element_line( color = "grey"),
        panel.grid.major.y = element_blank(), panel.border = element_blank(),
        panel.grid.minor = element_blank(), panel.background = element_blank(),
        axis.line = element_line(color = "black")) +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_vline(xintercept = 3.76, color = "red", linetype = 2, size = 1) +
  labs(subtitle = paste0("n = ", length(ucd$dx_date[ucd$sex == 1 & !is.na(ucd$dx_date)])),
       title = "Female",
       color = "Year") +
  ylab(paste0("Absolute number of cases (N=", length(ucd$dx_date[ucd$sex == 1 & !is.na(ucd$dx_date)]),")"))  +
  expand_limits(y = 0) +
  scale_x_continuous(name = "Month", breaks = month_breaks[1:12], labels =
                       month_names, expand = c(0,0)) +
  theme(plot.subtitle = element_text(hjust = 1), plot.caption =
          element_text(hjust = 1))

ggplotly(p2)


```

</br>
</br>

```{r plot3 , message=FALSE, echo=FALSE, warning=FALSE, error=FALSE}

p3<-summary_table %>%  filter(sex==0) %>% ggplot(aes(x=month, y=total_dx, color = as.factor(year))) +
  geom_line(size = 1) +
  theme(panel.grid.major.x = element_line( color = "grey"),
        panel.grid.major.y = element_blank(), panel.border = element_blank(),
        panel.grid.minor = element_blank(), panel.background = element_blank(),
        axis.line = element_line(color = "black")) +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_vline(xintercept = 3.76, color = "red", linetype = 2, size = 1) +
  labs(subtitle = paste0("n = ", length(ucd$dx_date[ucd$sex == 0 & !is.na(ucd$dx_date)])),
       title = "Male",
       color = "Year") +
  ylab(paste0("Absolute number of cases (N=", length(ucd$dx_date[ucd$sex == 0 & !is.na(ucd$dx_date)]),")"))  +
  expand_limits(y = 0) +
  scale_x_continuous(name = "Month", breaks = month_breaks[1:12], labels =
                       month_names, expand = c(0,0)) +
  theme(plot.subtitle = element_text(hjust = 1), plot.caption =
          element_text(hjust = 1))

ggplotly(p3)

```

</br>
</br>

```{r plot4,  message=FALSE, echo=FALSE, warning=FALSE, error=FALSE}

#############################################################################
#                             Admissions                                    #
#############################################################################

summary_table$sex[summary_table$sex == 0] <- "men"
summary_table$sex[summary_table$sex == 1] <- "women"


p1<- summary_table %>% select(year,month,sex,total_prim_visit) %>% 
  group_by(year,month)  %>% 
ggplot(aes(x=month, y=total_prim_visit, color = as.factor(year), linetype = as.factor(sex), fill = as.factor(sex))) +
  geom_line(size = 1) +
  theme(panel.grid.major.x = element_line( color = "grey"),
        panel.grid.major.y = element_blank(), panel.border = element_blank(),
        panel.grid.minor = element_blank(), panel.background = element_blank(),
        axis.line = element_line(color = "black")) +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_vline(xintercept = 3.76, color = "red", linetype = 2, size = 1) +
  labs(subtitle = paste0("n = ", sum(summary_table$total_prim_visit)),
       title = "Primary care admissions",
       color = "Year",
       lty = "Sex") +
  ylab(paste0("Absolute number of cases (N=", sum(summary_table$total_prim_visit),")"))  +
  expand_limits(y = 0) +
  scale_x_continuous(name = "Month", breaks = month_breaks[1:12], labels =
                       month_names, expand = c(0,0)) +
  theme(plot.subtitle = element_text(hjust = 1), plot.caption =
          element_text(hjust = 1))


ggplotly(p1)


```

`r if(sum(summary_table$total_prim_visit)==0){ warning("Warning: No information available on 'prim_visit'") }`

</br>
</br>

```{r plot5, message=FALSE, echo=FALSE, warning=FALSE, error=FALSE}

##


p2<-ggplot(summary_table, aes(x=month, y=total_hosp_visit, color = as.factor(year), linetype = as.factor(sex), fill = as.factor(sex))) +
  geom_line(size = 1) +
  theme(panel.grid.major.x = element_line( color = "grey"),
        panel.grid.major.y = element_blank(), panel.border = element_blank(),
        panel.grid.minor = element_blank(), panel.background = element_blank(),
        axis.line = element_line(color = "black")) +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_vline(xintercept = 3.76, color = "red", linetype = 2, size = 1) +
  labs(subtitle = paste0("n = ", sum(summary_table$total_hosp_visit)),
       title = "Hospital admissions",
       color = "Year",
       lty = "Sex") +
  ylab(paste0("Absolute number of cases (N=", sum(summary_table$total_hosp_visit),")"))  +
  expand_limits(y = 0) +
  scale_x_continuous(name = "Month", breaks = month_breaks[1:12], labels =
                       month_names, expand = c(0,0)) +
  theme(plot.subtitle = element_text(hjust = 1), plot.caption =
          element_text(hjust = 1))

ggplotly(p2)

```

`r if(sum(summary_table$total_hosp_visit)==0){ warning("Warning: No information available on 'hosp_visit'") }`

</br>
</br>

```{r plot6,  message=FALSE, echo=FALSE, warning=FALSE, error=FALSE}
p2<-ggplot(summary_table, aes(x=month, y=total_emer_visit, color = as.factor(year), linetype = as.factor(sex), fill = as.factor(sex))) +
  geom_line(size = 1) +
  theme(panel.grid.major.x = element_line( color = "grey"),
        panel.grid.major.y = element_blank(), panel.border = element_blank(),
        panel.grid.minor = element_blank(), panel.background = element_blank(),
        axis.line = element_line(color = "black")) +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_vline(xintercept = 3.76, color = "red", linetype = 2, size = 1) +
  labs(subtitle = paste0("n = ", sum(summary_table$total_emer_visit)),
       title = "Emergency admissions",
       color = "Year",
       lty = "Sex") +
  ylab(paste0("Absolute number of cases (N=", sum(summary_table$total_emer_visit),")"))  +
  expand_limits(y = 0) +
  scale_x_continuous(name = "Month", breaks = month_breaks[1:12], labels =
                       month_names, expand = c(0,0)) +
  theme(plot.subtitle = element_text(hjust = 1), plot.caption =
          element_text(hjust = 1))

ggplotly(p2)

```


`r if(sum(summary_table$total_emer_visit)==0){ warning("Warning: No information available on 'emer_visit'") }`
</br>
</br>

```{r plot7,  message=FALSE, echo=FALSE, warning=FALSE, error=FALSE}

#############################################################################
#                             Diagnosis of Anxiety                          #
#############################################################################
 

p1<-ggplot(summary_table, aes(x=month, y=total_anx, color = as.factor(year), linetype = as.factor(sex), fill = as.factor(sex))) +
  geom_line(size = 1) +
  theme(panel.grid.major.x = element_line( color = "grey"),
        panel.grid.major.y = element_blank(), panel.border = element_blank(),
        panel.grid.minor = element_blank(), panel.background = element_blank(),
        axis.line = element_line(color = "black")) +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_vline(xintercept = 3.76, color = "red", linetype = 2, size = 1) +
  labs(subtitle = paste0("n = ", sum(summary_table$total_anx)),
       title = "Diagnosis of Anxiety",
       color = "Year",
       lty = "Sex") +
  ylab(paste0("Absolute number of cases (N=", sum(summary_table$total_anx),")"))  +
  expand_limits(y = 0) +
  scale_x_continuous(name = "Month", breaks = month_breaks[1:12], labels =
                       month_names, expand = c(0,0)) +
  theme(plot.subtitle = element_text(hjust = 1), plot.caption =
          element_text(hjust = 1))

ggplotly(p1)


```

</br>
</br>

```{r plot8,  message=FALSE, echo=FALSE, warning=FALSE, error=FALSE}

#############################################################################
#                             Diagnosis of Anxiety                          #
#############################################################################


p1<-ggplot(summary_table, aes(x=month, y=total_dep, color = as.factor(year), linetype = as.factor(sex), fill = as.factor(sex))) +
  geom_line(size = 1) +
  theme(panel.grid.major.x = element_line( color = "grey"),
        panel.grid.major.y = element_blank(), panel.border = element_blank(),
        panel.grid.minor = element_blank(), panel.background = element_blank(),
        axis.line = element_line(color = "black")) +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_vline(xintercept = 3.76, color = "red", linetype = 2, size = 1) +
  labs(subtitle = paste0("n = ", sum(summary_table$total_dep)),
       title = "Diagnosis of Depression",
       color = "Year",
       lty = "Sex") +
  ylab(paste0("Absolute number of cases (N=", sum(summary_table$total_dep),")"))  +
  expand_limits(y = 0) +
  scale_x_continuous(name = "Month", breaks = month_breaks[1:12], labels =
                       month_names, expand = c(0,0)) +
  theme(plot.subtitle = element_text(hjust = 1), plot.caption =
          element_text(hjust = 1))

ggplotly(p1)


```

</br>
</br>

### Summary table with total monthly cases by indicator ### 

```{r plottable , message=FALSE, echo=FALSE, warning=FALSE, error=FALSE}

summary_table$sex[summary_table$sex == "men"] <- 0
summary_table$sex[summary_table$sex == "women"] <- 1

datatable(summary_table, class = 'cell-border stripe')

```

</br>
</br>

```{r}



ucd$year <- as.numeric(strftime(ucd$dx_date, "%Y"))    # Create year column
ucd$month <- as.numeric(strftime(ucd$dx_date, "%m"))

dataAva <- sum(is.na(ucd$year)) < length(ucd$year)

if(dataAva){
  ucd <- ucd %>%  mutate(year_month = as.Date(paste0(year,'-',month,"-01"))) %>% filter(year_month >= "2017-01-01")



 
init_year <- min(ucd$year,na.rm = TRUE)
end_year <- max(ucd$year,na.rm = TRUE)


  #In the next step, we can apply the aggregate function to convert our daily data to monthly data:
new_dx =data.frame(aggregate(dep ~ month + year, ucd , FUN = sum), 
                    aggregate(anx ~ month + year, ucd, FUN = sum),
                    aggregate(cbind(count = patient_id) ~ month + year, ucd, FUN = function(x){NROW(x)}))

## TODO ATENTO SI ES MAS DE 2020
# Dx depression per month
ts_new_dep <- ts(new_dx$dep, start=c(init_year, 1), end=c(end_year, 12), frequency=12)
ts_new_anx <- ts(new_dx$anx, start=c(init_year, 1), end=c(end_year, 12), frequency=12)

ts.plot(ts_new_dep, ts_new_anx, gpars = list(col = c("black", "red"),
                                                  xlab="month of diagnosis",
                                                  ylab="n", lty=c(1:2)),
             main = "New depression/anxiety diagnosis") 


legend("topright", legend = c("Depression", "Anxiety"), col = 1:2, lty=c(1:2))

}



```
`r if(!dataAva){ warning("Warning: No information available on 'dx_date'") }`

</br>
</br>

```{r}

if(dataAva){

init_year <- year(min(ucd$year_month,na.rm = TRUE ))
end_year <- year(max(ucd$year_month,na.rm = TRUE))
diff_years <- (end_year - init_year) + 1 

dx_dep_mo <- data.frame(mo=rep(1:12,diff_years),yr=factor(rep(init_year:end_year,each=12)),
                  aggregate(ucd$dep, by = list(ucd$year_month), FUN = sum),
                  aggregate(ucd$patient_id, by = list(ucd$year_month), FUN = length))

dx_dep_mo$prop_dep_mo=dx_dep_mo$x*100/dx_dep_mo$x.1


# plot

p <- ggplot(dx_dep_mo,aes(x=mo,y=prop_dep_mo, col=yr,group=yr))+
  geom_line(size=0.5, aes(color=yr))+ geom_point(size=3)+
  labs(x="Month of diagnosis", y="Percentage of new depressions", 
       title = "New diagnoses of depression among patients with ill-mental health",color = "Year")  + ylim(0, 100) +  
  scale_x_continuous(breaks=c(1:12),
       labels=c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))+
  theme(panel.grid.major.x = element_line( color = "grey"),
        panel.grid.major.y = element_blank(), panel.border = element_blank(),
        panel.grid.minor = element_blank(), panel.background = element_blank(),
        axis.line = element_line(color = "black")) 


ggplotly(p)

}

```

`r if(!dataAva){ warning("Warning: No information available on 'dx_date'") }`

</br>
</br>

```{r}
# anxiety per month
if(dataAva){

col1 <- aggregate(ucd$anx, by = list(ucd$year_month), FUN = sum)
names(col1) <- c("fecha","x")

dx_anx_mo=data.frame(mo=rep(1:12,diff_years),yr=factor(rep(init_year:end_year,each=12)),
                  aggregate(ucd$anx, by = list(ucd$year_month), FUN = sum),
                  aggregate(ucd$patient_id, by = list(ucd$year_month), FUN = length))

dx_anx_mo$prop_anx_mo=dx_anx_mo$x*100/dx_anx_mo$x.1


# plot

p2 <- ggplot(dx_anx_mo,aes(x=mo,y=prop_anx_mo, col=yr,group=yr))+
  geom_line(size=0.5, aes(color=yr))+ geom_point(size=3)+
  labs(x="Month of diagnosis", y="Percentage of new anxieties", 
       title = "New diagnoses of anxiety among patients with ill-mental health",color = "Year")  + ylim(0, 100) +  
  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9,10,11,12),
                     labels=c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))+
  theme(panel.grid.major.x = element_line( color = "grey"),
        panel.grid.major.y = element_blank(), panel.border = element_blank(),
        panel.grid.minor = element_blank(), panel.background = element_blank(),
        axis.line = element_line(color = "black")) 

ggplotly(p2)
}



```
`r if(!dataAva){ warning("Warning: No information available on 'dx_date'") }`

```{r write_aggregated,  message=FALSE, echo=FALSE, warning=FALSE, error=FALSE}

## add country

summary_table_aggregated$country <- unique(ucd$country_cd)[1]
write.csv(summary_table_aggregated, file = "aggregated_outputs.csv", row.names = FALSE)


```

